Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}

### 2020-06-20 revision_2/paper.tex
clean up to output mse results for revision_2, minor edits to paper.:

\response{The advantages and disadvantages of the method are detailed throughout the paper using simulated data and real data. One conceptual advantage is that inferences and predictions are less sensitive to features of the data not captured by
the conditioning statistics. Choosing statistics targeting main features of interest allows for more targeted inference on these features. The analysis can help to better understand other features, such as outliers, not capture by the conditioning statistics.} 
\response{The examples in the paper suggest advantages in situations where outliers are concern and there is significant prior information for the non-outlying portion of the model that is not outweighed by the data. The main disadvantage over traditional robust estimate techniques are mainly computational. In Section ~\ref{BayesLinMod} we detail a data-augmented MCMC algorithm to fit the models proposed in this paper. This requires an additional computational step for each iteration of the chain. Details of the additional burden are given and one must weigh the advantages of these methods with this additional burden. Since it is typically that the restricted likelihood posterior converges to the same expected value as the conditioning statistics as the sample size grows, it may not be practical to implement our method in all situations.}

The remainder of the paper is as follows: Section~\ref{restrictedlikelihood} introduces the Bayesian restricted likelihood, provides context with previous work, and demonstrates some advantages of the methods on simple examples. Section~\ref{BayesLinMod} details an MCMC algorithm to apply the method to Bayesian linear models. This computational strategy is a major contribution to the work, providing an approach to apply the method on realistic examples. Many of the the technical proofs are in the Appendix \ref{sec:appendix} with \texttt{R} code available from the authors. Sections \ref{simData} and \ref{RealData} illustrate the method with simulated data and a real insurance industry data set containing many outliers with a novel twist on model evaluation. A discussion (Section~\ref{Conclusions}) provides some final commentary on the new method. \response{ An R package \texttt{brlm} to implement our methods is available at \texttt{github.com/jrlewi/brlm}. Additionally all data and code for the examples in this paper are available at \texttt{https://github.com/jrlewi/brlm\char`_paper/revision\char`_2}.}

@@

\subsection{The Bayesian linear model}
We focus on the use of restricted likelihood for the Bayesian linear
model with a standard formulation: 
\begin{eqnarray}
\label{LinearModel}
\bth&=&(\bbeta,\sigma^2) \sim  \pi(\bth) 
\nonumber\\
y_i  & =  & x_i^\top \bbeta + \epsilon_i , \mbox{ for } i = 1, \ldots, n 
\end{eqnarray}
where $x_i$ and $\bbeta \in \mathbb{R}^p$, $\sigma^2 \in \mathbb{R}^+$, 
and the $\epsilon_i$ are independent draws from a distribution with center $0$ and scale $\sigma$. $X$ denotes the design matrix whose rows are  $x_i^\top$. For the restricted likelihood model,  conditioning statistics are assumed to be of the form $T(\by) = (\bb(X, \by), s(X, \by))$ where $\bb(X, \by)= (b_1(X,\by), \dots,b_p(X,\by))^\top\in \mathbb R^{p}$ is an estimator for the regression coefficients and $s(X, \by)\in \{0\} \cup {\mathbb R}^+$ is an estimator of the scale. Throughout, observed data and summary statistic is denoted by $\by_{obs}$ and $T(\by_{obs})=(\bb(X, \by_{obs}), s(X, \by_{obs}))$, respectively. 
Several conditions are imposed on the model and statistic to ensure validity of the MCMC algorithm:
\begin{itemize}
\labitem{C1}{fullRank} The $n \times p$ design matrix, $X$, whose $i^{th}$ row is $x_i^\top$, 
is of full column rank.  
\labitem{C2}{supReal} The $\epsilon_i$ are a random sample from some distribution which has a density with 
respect to Lebesgue measure on the real line and for which the support is the real line.  
\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
\end{itemize}
Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
\textit{regression} and \textit{scale equivariance},
respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
and \textit{scale equivariance}. 
Many estimators satisfy the above properties, including several traditional simultaneous M-estimators \citep{huber2009, maronna2006} for which the \texttt{R} package \texttt{brlm} (\texttt{github.com/jrlewi/brlm}) is available to implement the MCMC described here. \response{These M-estimators satisfy \ref{asb}-\ref{as} since they are optimizers of (almost surely) continuous and differentiable objective functions. Constraints  \ref{regEq}-\ref{scaleEq2Reg} are often satisfied by location and scale estimators but should be checked on a case by case basis.}  More software development is required to extend the MCMC implementation beyond the M-estimators discussed here. The current version of the \texttt{R} package also implements the direct computational methods described in \cite{lewis2014}. These methods are effective in lower dimensional problems and were used in both examples in Section \ref{illustrations}.

@@

\response{Finally, it is clear the estimators themselves must be computed for every iteration of the Markov Chain. We have found this burden to be marginal in relation with respect to computing the needed Jacobian. In the simulations and real data analyses presented below, we will see that the additional burden is often of added value over traditional robust regression when substantial prior information is available that is not swamped by current data.}

@@

\section{Simulated Data}
\label{simData}
We study the performance of restricted likelihood methods in two simulation settings. The first is a hierarchical setting. The second is a variable selection setting where there are several potential covariates where only a few have non-zero effect sizes.

@@

In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
with $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ and the error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.8$ and $\epsilon \sim \text{Half-Normal}(0,25\sigma^{2})$ with probability $0.2$ (i.e., there is a relatively large amount of one-sided outlier contamination). The components of $x = (x_{1}, x_{2}, x_{3})$ are correlated with $x_1 \sim N(0,1)$ and $x_{j} = x_{1} + \eta_{j}$ with $\eta_{j} \sim N(0, 4)$ for $j = 2,3$. This results in a theoretical correlation of $1/sqrt(5) \approx 0.44$ between $x_{1}$ and both $x_{j}, j = 2, 3$. The model used for fitting contains an additional 27 covariates, some of which are also correlated with $x_{1}$, $x_{2}$, and $x_{3}$. Specifically the fitting model is $y = \beta^{\top} x + \beta^{*\top} x^{*} + \epsilon$ where $x^{*}$ and $\beta^{*}$ are 27 dimensional vectors of extra covariates and slope parameters. Of these 27 covariates, 21 are generated independently from standard normal distributions. Of the remaining 6, two each are generated by adding standard normal noise to $x_{1}$, $x_{2}$, and $x_{3}$. This represents a common situation where several covariates with various levels of correlation amongst them are available for fitting, but only a few govern the data generating mechanism.

For the simulation, $K = 30$ data sets (including the additional covariates) of size $n = 500$ are generated from the true model with true values $\bbeta = (1,1,1)^{\top}$ and $\sigma^{2} = 2$. We fit the model including all 30 covariates and consider the following methods for the fit 1) classical robust regression with Tukey's estimator of location and Huber's estimator of scale, 2) the corresponding restricted likelihood version 3) A heavy-tailed Bayesian model with a Student-t likelihood with $\nu = 5$ degrees of freedom.  For the Bayesian models we take $\bbeta_{all}\sim N_{20}(\mb{0}, \sigma^{2}_{\beta}I)$ with $\bbeta_{all} = (\bbeta, \bbeta^{*})^{\top}$ and $\sigma^{2}\sim IG(5,8)$ under the restricted model and  $\sigma^{2}\sim IG(5,\frac{\nu-2}{\nu}8)$ under the Student-t model.  For each data set, we fit the models for $ \sigma_{\beta} = 0.4, 0.6, 0.8,\dots, 1.4$. The acceptance rates for the restricted likelihood MCMC data-augmentation step range from $0.3$ to $0.36$ across all the data sets and values of $\sigma_{\beta}$. To compare performance we first consider the $MSE = (||\bbeta - \hat\bbeta||^{2} + ||\hat\bbeta^{*}||^{2})/30$ for each simulation where $\hat\bbeta$ and $\hat\bbeta^{*}$ are point estimates for the fitted model. For the Bayesian models, we use posterior means. The average MSE plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{mseSimMany}. The classical fit is labeled `rlm` and is the same for each value of the prior standard deviation $\sigma^{2}_{\beta}$. We see for most values of the the prior standard deviation, the Bayesian models (`restricted' and `t') outperform the classical fit. The correlation amongst the covariates causes a certain level of confounding and the prior shrinkage helps to improve estimation. However, too much shrinkage can be detrimental as demonstrated for $\sigma_{\beta} = 0.4$. While this will help for estimation of $\bbeta^{*} = 0$, the estimation of the active parameters $\bbeta$ can be hindered. The $t$ model seems more sensitive to this effect than the restricted model. The restricted model also has an additional advantage when it comes to prediction of the non-outlying data. To see this, for each simulation we consider the mean negative log-likelihood of the non-outlying data: $MNLL = -\frac{1}{N} \sum \log f(y_{i} |  \hat\bbeta, \hat\bbeta^{*}, \hat\sigma)$ where $f$ is the assumed likelihood and the average is taken over the $N$ non-outlying points $y_{i}$. For the classical and restricted fits, $f$ is the normal likelihood and for the `t' it is the heavy-tailed Student-t likelihood. The average MNLL plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{negllSimMany}. First, the restricted version has a small but consistent improvement over the classical method. Second, it is clear that that the heavy-tailed model suffers when trying to predict the non-outlying data since it assumes the entire data generating mechanism is heavy-tailed.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim_many_p.png}
\caption{Average MSE plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma^{2}_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma^{2}_{\beta}$.}
\label{mseSimMany}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{negll_sim_many_p.png}
\caption{Average MNLL plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma^{2}_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma^{2}_{\beta}$.}
\label{negllSimMany}
\end{figure}

@@

Models are fit to four states labelled State 2, 15, 27, and 36, with $n = 222, 40, 117,$ and $46$, representing a range of sample sizes. Fitting is done on $K = 50$ training samples with training sample sizes taken to be $0.25n$ and $0.50n$. Holdout evaluation is done on the remaining (`Type 1') samples. The acceptance rates for  the data augmentation step,\response{for all but one training set}, range from  \response{$0.10$ to $0.8$} across the states, repetitions, and two versions of the model. \response{The exception was a single training set from State 15 resulting in an usually small acceptance rate under Tukey's version. This case didn't effect the overall results of the simulations but emphasizes the need to check convergence on a case by case basis.} The average $TLM_b(A)$ over the $K = 50$ training/holdout samples for the four states and seven methods are shown in Figure \ref{fig:tlm} where the base model is the Student-t model and $\alpha = 0.3$. Similar results are observed for other base models. The error bars are plus/minus one standard deviation of the average $TLM_b(A)$ over the $K = 50$  training/holdout samples. It is clear that the normal Bayesian model used as the full model (Normal) and the classical ordinary least squares fits (OLS) have poor performance due to the significant amount of outlier contamination in the data. In comparing our restricted methods to their corresponding classical methods, there is small, but consistent improvement across the states and training sample size. Additionally, variance reduction for the Bayesian versions is evident, especially in State 15, highlighted by the smaller error bars. For state 2, the largest state with $n = 222$, the restricted and classical robust methods have similar performance especially for larger training sample size. This reflects the diminishing effect of the prior as the sample size grows. Notably, the Student-t model performs poorly in comparison for this state. The predictive distribution explicitly accounts for heavy-tailed values, resulting in poorer predictions of the `good' data (i.e., the Type 1 agencies). Likewise, for State 27, another larger state, the Student-t model is outperformed by our restricted methods.   For the other states (State 15 and 36), the Student-t performs better to our restricted methods for smaller training sample size (25\% of the sample). However, this advantage goes away for the larger training sample size (50\% of the sample). Intuitively, as more data is available for fitting, more outliers appear and the heavy-tailed model compensates for them by assuming they come from the tails of the model; an assumption which is detrimental for prediction. Comparisons of the models depend on $\alpha$ as seen in Figure \ref{fig:tlmbyAlpha} which shows results for different $\alpha$ for training sample size $0.5n$. For smaller $\alpha$ (in this case $\alpha = 0.1$), many outliers are left untrimmed resulting in lower TLM  for all methods and noticeably larger standard deviation for the classical robust methods and our restricted likelihood. Larger values of $\alpha$ ensure that the predictive performance assessment excludes the majority of outliers. The proportion of $0$ counts in the data is roughly $0.14$, suggesting that $\alpha$ should be at least this large.

@@

\subsection{Practical Considerations for Using the Restricted Likelihood}
We offer the following guidelines and recommendations for using these methods in practice. 
\begin{itemize}
\item We have shown the method to be useful in outlier prone situations where good prior information can be encoded for the `good' part of the model. The additional computation burden may be too much when this is not the case. Traditional robust estimation techniques are often quicker and can provide comparable results.
\item Currently, we recommend conditioning on classical robust estimators when using these methods. Examples include Huber's and Tukey's estimators as outlined in the paper. We have found default choices (e.g. obtaining 95\% efficiency under normally distributed data) for the tuning parameters to work well in many situations. However, depending on the amount of contamination, some adjustment may be necessary. 
\item The likelihood itself can indeed take many forms and, like the conditioning statistic, can depend on the application. We have developed methods for fitting these models under a large class of conditioning statistics. In outlier prone situations we have had success using a normal likelihood model, with the conditioning intended to reduce the sensitivity of the analysis to the outliers. 
\item Comparisons amongst various choices of conditioning statistics and models can be made via standard cross-validation techniques like those done in this paper where models are fit to several training and holdout sets. One should take care in the method of evaluation. When the goal is to obtain good estimation/prediction for the non-outlying data, evaluation should take place on non-outlying data or the metric should be adjusted to discount outlying data ins some way (such as the trimmed-log-marginal we introduced TLM).
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{M-estimators}
\label{sec:Mestimators}
%%%%%%%%%%%%%%%%%%%%%%%%
M-estimators offer a natural choice for conditioning in restricted likelihood settings and they can be readily applied when conditioning on these estimators using the method described in this paper. This section gives a brief review. General estimating equations for the simultaneous M-estimators used in this paper appear in equation \ref{Mest} of the main text. Taking $\psi = \rho'$, the two $\rho$ functions used in this paper are: 
\begin{equation}
\label{Huber}
\rho(x)=
\begin{cases}
x^2, & \  |x|\leq k \\
2k|x|-k^2,&\ |x|\geq k 
\end{cases}
\end{equation}
and 
\begin{equation}
\label{Tukey}
\rho(x)=
\begin{cases}
1-[1-(x/k)^2]^3, & \  |x|\leq k \\
1, &\ |x|\geq k .
\end{cases}
\end{equation}
The first of these is known as Huber's $\rho$. It offers a compromise to squared error loss, with the loss becoming linear for absolute residuals larger than $k$. \cite{huber2009} provide a theoretical justification of this choice based the \textit{least informative distribution} in a particular class of contaminated normal distributions. The $\rho$ in \eqref{Tukey} is known as Tukey's bisquare. Tukey's $\psi = \rho'$ is \textit{re-descending} as it returns to zero outside of $(-k,k)$. As a result, Tukey's estimator downweights large residuals more than does Huber's estimator. Huber's estimator is akin to Winsorising whereas Tukey's is akin to complete trimming. Note that $k$ is a tuning constant which must be chosen. We only consider `default' tuning parameters in this paper which have 95\% asymptotic efficiency under a normal distribution. More detailed results can be found in both \cite{huber2009} and \cite{maronna2006}. These values are $k=1.345$ for Huber's $\rho$ and $k=4.685$ for Tukey's.

For the $\chi$ function, \cite{huber1964} proposed the choice 
\begin{equation}
\label{HuberScale}
\chi(x)=
\begin{cases}
x^2-\delta, &\ |x|\leq k\\
k^2-\delta, &\ |x|> k .
\end{cases}
\end{equation}
For consistency under the normal distribution the tuning parameter is set to $\delta\approx 0.71016$, taking $k = 1.345$ as above for Huber's $\psi$ is
referred to as Huber's `Proposal 2'. This is estimator used throughout this this paper.

>>>>

\response{The advantages and disadvantages of the method are detailed throughout the paper using simulated data and real data. One conceptual advantage is that inferences and predictions are less sensitive to features of the data not captured by
the conditioning statistics. Choosing statistics targeting main features of interest allows for more targeted inference on these features. The analysis can help to better understand other features, such as outliers, not captured by the conditioning statistics.} 
\response{The examples in the paper suggest advantages in situations where outliers are concern and there is significant prior information for the non-outlying portion of the model that is not outweighed by the data. The main disadvantage over traditional robust estimation techniques are mainly computational. In Section ~\ref{BayesLinMod} we detail a data-augmented MCMC algorithm to fit the models proposed in this paper. This requires an additional computational step for each iteration of the chain. Details of the additional burden are given and one must weigh the advantages of these methods with this additional burden. Since the restricted likelihood posterior typically converges to the same expected value as the conditioning statistics, it may not be practical to implement our method in all situations.}

The remainder of the paper is as follows: Section~\ref{restrictedlikelihood} introduces the Bayesian restricted likelihood, provides context with previous work, and demonstrates some advantages of the methods on simple examples. Section~\ref{BayesLinMod} details an MCMC algorithm to apply the method to Bayesian linear models. This computational strategy is a major contribution to the work, providing an approach to apply the method on realistic examples. Many of the the technical proofs are in the Appendix \ref{sec:appendix} with \texttt{R} code available from the authors. Sections \ref{simData} and \ref{RealData} illustrate the method with simulated data and a real insurance industry data set containing many outliers with a novel twist on model evaluation. A discussion (Section~\ref{Conclusions}) provides some final commentary on the new method. \response{ An R package \texttt{brlm} to implement our methods is available at \texttt{github.com/jrlewi/brlm}. Additionally all data and code for the examples in this paper are available at \texttt{https://github.com/jrlewi/brlm\char`_paper}.}

@@

\subsection{The Bayesian linear model}
We focus on the use of restricted likelihood for the Bayesian linear
model with a standard formulation: 
\begin{eqnarray}
\label{LinearModel}
\bth&=&(\bbeta,\sigma^2) \sim  \pi(\bth) 
\nonumber\\
y_i  & =  & x_i^\top \bbeta + \epsilon_i , \mbox{ for } i = 1, \ldots, n 
\end{eqnarray}
where $x_i$ and $\bbeta \in \mathbb{R}^p$, $\sigma^2 \in \mathbb{R}^+$, 
and the $\epsilon_i$ are independent draws from a distribution with center $0$ and scale $\sigma$. $X$ denotes the design matrix whose rows are  $x_i^\top$. For the restricted likelihood model,  conditioning statistics are assumed to be of the form $T(\by) = (\bb(X, \by), s(X, \by))$ where $\bb(X, \by)= (b_1(X,\by), \dots,b_p(X,\by))^\top\in \mathbb R^{p}$ is an estimator for the regression coefficients and $s(X, \by)\in \{0\} \cup {\mathbb R}^+$ is an estimator of the scale. Throughout, observed data and summary statistic is denoted by $\by_{obs}$ and $T(\by_{obs})=(\bb(X, \by_{obs}), s(X, \by_{obs}))$, respectively. 
Several conditions are imposed on the model and statistic to ensure validity of the MCMC algorithm:
\begin{itemize}
\labitem{C1}{fullRank} The $n \times p$ design matrix, $X$, whose $i^{th}$ row is $x_i^\top$, 
is of full column rank.  
\labitem{C2}{supReal} The $\epsilon_i$ are a random sample from some distribution which has a density with 
respect to Lebesgue measure on the real line and for which the support is the real line.  
\labitem{C3}{asb}$\bb(X,\by)$ is almost surely continuous and differentiable with respect to $\by$.  
\labitem{C4}{as} $s(X,\by)$ is almost surely positive, continuous, and differentiable with respect to $\by$.  
\labitem{C5}{regEq} $\bb(X,\by+X\bv)=\bb(X,\by)+\bv \ \ \text{for  all}\ \bv\in\mathbb{R}^p$. 
\labitem{C6}{scaleEqReg} $\bb(X,a\by)=a\bb(X,\by)\ \ \ \text{for all constants } a$.  
\labitem{C7}{regIn} $s(X,\by+X\bv)=s(X,\by) \ \ \text{for all}\ \bv\in\mathbb{R}^p$.  
\labitem{C8}{scaleEq2Reg} $s(X, a\by)=|a|s(X,\by) \ \ \text{for all constants } a$.  
\end{itemize}
Properties \ref{regEq} and \ref{scaleEqReg} of $\bb$ are called
\textit{regression} and \textit{scale equivariance},
respectively.  Properties \ref{regIn} and \ref{scaleEq2Reg} of $s$ are called \textit{regression invariance}
and \textit{scale equivariance}. 
Many estimators satisfy the above properties, including several traditional simultaneous M-estimators \citep{huber2009, maronna2006} for which the \texttt{R} package \texttt{brlm} (\texttt{github.com/jrlewi/brlm}) is available to implement the MCMC described here. \response{These M-estimators satisfy \ref{asb}-\ref{as} since they are optimizers of (almost surely) continuous and differentiable objective functions. Constraints  \ref{regEq}-\ref{scaleEq2Reg} are often satisfied by location and scale estimators but should be checked on a case-by-case basis.}  More software development is required to extend the MCMC implementation beyond the M-estimators discussed here. The current version of the \texttt{R} package also implements the direct computational methods described in \cite{lewis2014}. These methods are effective in lower dimensional problems and were used in both examples in Section \ref{illustrations}.

@@

\response{Finally, it is clear the estimators themselves must be computed for every iteration of the Markov Chain. We have found this burden to be marginal in comparison to computing the needed Jacobian. In the simulations and real data analyses presented below, we will see that the additional computational expense needed to fit the Bayesian model is often worthwhile, leading to better performance compared to traditional, non-Bayesian robust regression estimators.  This is most evident when substantive prior information is available and information in the data is limited.}

@@

\section{Simulated Data}
\label{simData}
We study the performance of restricted likelihood methods in two simulation settings. The first is a hierarchical setting. The second is a variable selection setting where there are several potential covariates but only a few have non-zero effect sizes.

@@

In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
with $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ and the error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.8$ and $\epsilon \sim \text{Half-Normal}(0,25\sigma^{2})$ with probability $0.2$ (i.e., there is a relatively large amount of one-sided outlier contamination). The components of $x = (x_{1}, x_{2}, x_{3})$ are correlated with $x_1 \sim N(0,1)$ and $x_{j} = x_{1} + \eta_{j}$ with $\eta_{j} \sim N(0, 4)$ for $j = 2,3$. This results in a theoretical correlation of $1/\sqrt{5} \approx 0.44$ between $x_{1}$ and both $x_{j}, j = 2, 3$. The model used for fitting contains an additional 27 covariates, some of which are also correlated with $x_{1}$, $x_{2}$, and $x_{3}$. Specifically the fitting model is $y = \beta^{\top} x + \beta^{*\top} x^{*} + \epsilon$ where $x^{*}$ and $\beta^{*}$ are 27 dimensional vectors of extra covariates and slope parameters. Of these 27 covariates, 21 are generated independently from standard normal distributions. Of the remaining 6, two each are generated by adding standard normal noise to $x_{1}$, $x_{2}$, and $x_{3}$. This represents a common situation where several covariates with various levels of correlation amongst them are available for fitting, but only a few govern the data generating mechanism.

For the simulation, $K = 30$ data sets (including the additional covariates) of size $n = 500$ are generated from the true model with true values $\bbeta = (1,1,1)^{\top}$ and $\sigma^{2} = 2$. We fit the model including all 30 covariates and consider the following methods for the fit 1) classical robust regression with Tukey's estimator of location and Huber's estimator of scale, 2) the corresponding restricted likelihood version 3) A heavy-tailed Bayesian model with a Student-t likelihood with $\nu = 5$ degrees of freedom.  For the Bayesian models we take $\bbeta_{all}\sim N_{30}(\mb{0}, \sigma^{2}_{\beta}I)$ with $\bbeta_{all} = (\bbeta, \bbeta^{*})^{\top}$ and $\sigma^{2}\sim IG(5,8)$ under the restricted model and  $\sigma^{2}\sim IG(5,\frac{\nu-2}{\nu}8)$ under the Student-t model.  For each data set, we fit the models for $ \sigma_{\beta} = 0.4, 0.6, 0.8,\dots, 1.4$. The acceptance rates for the restricted likelihood MCMC data-augmentation step range from $0.3$ to $0.36$ across all the data sets and values of $\sigma_{\beta}$. To compare performance we first consider the $MSE = (||\bbeta - \hat\bbeta||^{2} + ||\hat\bbeta^{*}||^{2})/30$ for each simulation where $\hat\bbeta$ and $\hat\bbeta^{*}$ are point estimates for the fitted model. For the Bayesian models, we use posterior means. The average MSE plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{mseSimMany}. The classical fit is labeled `rlm' and is the same for each value of the prior standard deviation $\sigma^{2}_{\beta}$. We see for most values of the the prior standard deviation, the Bayesian models (`restricted' and `t') outperform the classical fit. The correlation amongst the covariates causes a certain level of confounding and the prior shrinkage helps to improve estimation. However, too much shrinkage can be detrimental as demonstrated for $\sigma_{\beta} = 0.4$. While this will help for estimation of $\bbeta^{*} = 0$, the estimation of the active parameters $\bbeta$ can be hindered. The $t$ model seems more sensitive to this effect than the restricted model. The restricted model also has an additional advantage when it comes to prediction of the non-outlying data. To see this, for each simulation we consider the mean negative log-likelihood of the non-outlying data: $MNLL = -\frac{1}{N} \sum \log f(y_{i} |  \hat\bbeta, \hat\bbeta^{*}, \hat\sigma)$ where $f$ is the assumed likelihood and the average is taken over the $N$ non-outlying points $y_{i}$. For the classical and restricted fits, $f$ is the normal likelihood and for the `t' it is the heavy-tailed Student-t likelihood. The average MNLL plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{negllSimMany}. First, the restricted version has a small but consistent improvement over the classical method. Second, it is clear that that the heavy-tailed model suffers when trying to predict the non-outlying data since it assumes the entire data generating mechanism is heavy-tailed.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim_many_p.png}
\caption{Average MSE plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma_{\beta}$.}
\label{mseSimMany}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{negll_sim_many_p.png}
\caption{Average MNLL plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma_{\beta}$.}
\label{negllSimMany}
\end{figure}

@@

Models are fit to four states labelled State 2, 15, 27, and 36, with $n = 222, 40, 117,$ and $46$, representing a range of sample sizes. Fitting is done on $K = 50$ training samples with training sample sizes taken to be $0.25n$ and $0.50n$. Holdout evaluation is done on the remaining (`Type 1') samples. The acceptance rates for  the data augmentation step, \response{for all but one training set}, range from  \response{$0.10$ to $0.8$} across the states, repetitions, and two versions of the model. \response{The exception was a single training set from State 15 resulting in an usually small acceptance rate under Tukey's version. This case didn't effect the overall results of the simulations but emphasizes the need to check convergence on a case-by-case basis.} The average $TLM_b(A)$ over the $K = 50$ training/holdout samples for the four states and seven methods are shown in Figure \ref{fig:tlm} where the base model is the Student-t model and $\alpha = 0.3$. Similar results are observed for other base models. The error bars are plus/minus one standard deviation of the average $TLM_b(A)$ over the $K = 50$  training/holdout samples. It is clear that the normal Bayesian model used as the full model (Normal) and the classical ordinary least squares fits (OLS) have poor performance due to the significant amount of outlier contamination in the data. In comparing our restricted methods to their corresponding classical methods, there is small, but consistent improvement across the states and training sample size. Additionally, variance reduction for the Bayesian versions is evident, especially in State 15, highlighted by the smaller error bars. For state 2, the largest state with $n = 222$, the restricted and classical robust methods have similar performance especially for larger training sample size. This reflects the diminishing effect of the prior as the sample size grows. Notably, the Student-t model performs poorly in comparison for this state. The predictive distribution explicitly accounts for heavy-tailed values, resulting in poorer predictions of the `good' data (i.e., the Type 1 agencies). Likewise, for State 27, another larger state, the Student-t model is outperformed by our restricted methods.   For the other states (State 15 and 36), the Student-t performs better to our restricted methods for smaller training sample size (25\% of the sample). However, this advantage goes away for the larger training sample size (50\% of the sample). Intuitively, as more data is available for fitting, more outliers appear and the heavy-tailed model compensates for them by assuming they come from the tails of the model; an assumption which is detrimental for prediction. Comparisons of the models depend on $\alpha$ as seen in Figure \ref{fig:tlmbyAlpha} which shows results for different $\alpha$ for training sample size $0.5n$. For smaller $\alpha$ (in this case $\alpha = 0.1$), many outliers are left untrimmed resulting in lower TLM  for all methods and noticeably larger standard deviation for the classical robust methods and our restricted likelihood. Larger values of $\alpha$ ensure that the predictive performance assessment excludes the majority of outliers. The proportion of $0$ counts in the data is roughly $0.14$, suggesting that $\alpha$ should be at least this large.

@@

\subsection{Practical Considerations for Using the Restricted Likelihood}
We offer the following guidelines and recommendations for using these methods in practice. 
\begin{itemize}
\item We have shown the method to be useful in outlier prone situations where good prior information can be encoded for the `good' part of the model. The additional computation burden may be too much when this is not the case. Traditional robust estimation techniques are often quicker and can provide comparable results.
\item Currently, we recommend conditioning on classical robust estimators when using these methods. Examples include Huber's and Tukey's estimators as outlined in the paper. We have found default choices (e.g. obtaining 95\% efficiency under normally distributed data) for the tuning parameters to work well in many situations. However, depending on the amount of contamination, some adjustment may be necessary. 
\item The likelihood itself can indeed take many forms and, like the conditioning statistic, can depend on the application. We have developed methods for fitting these models under a large class of conditioning statistics. In outlier prone situations we have had success using a normal likelihood model, with the conditioning intended to reduce the sensitivity of the analysis to the outliers. 
\item Comparisons amongst various choices of conditioning statistics and models can be made via standard cross-validation techniques like those done in this paper where models are fit to several training and holdout sets. One should take care in the method of evaluation. When the goal is to obtain good estimation/prediction for the non-outlying data, evaluation should take place on non-outlying data or the metric should be adjusted to discount outlying data in some way (such as the trimmed-log-marginal we introduced).
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{M-estimators}
\label{sec:Mestimators}
%%%%%%%%%%%%%%%%%%%%%%%%
M-estimators offer a natural choice for conditioning in restricted likelihood settings and the methods described in this paper can be readily applied when conditioning on these estimators. This section gives a brief review. General estimating equations for the simultaneous M-estimators used in this paper appear in equation \ref{Mest} of the main text. Taking $\psi = \rho'$, the two $\rho$ functions used in this paper are: 
\begin{equation}
\label{Huber}
\rho(x)=
\begin{cases}
x^2, & \  |x|\leq k \\
2k|x|-k^2,&\ |x|\geq k 
\end{cases}
\end{equation}
and 
\begin{equation}
\label{Tukey}
\rho(x)=
\begin{cases}
1-[1-(x/k)^2]^3, & \  |x|\leq k \\
1, &\ |x|\geq k .
\end{cases}
\end{equation}
The first of these is known as Huber's $\rho$. It offers a compromise to squared error loss, with the loss becoming linear for absolute residuals larger than $k$. \cite{huber2009} provide a theoretical justification of this choice based the \textit{least informative distribution} in a particular class of contaminated normal distributions. The $\rho$ in \eqref{Tukey} is known as Tukey's bisquare. Tukey's $\psi = \rho'$ is \textit{re-descending} as it returns to zero outside of $(-k,k)$. As a result, Tukey's estimator downweights large residuals more than does Huber's estimator. Huber's estimator is akin to Winsorising whereas Tukey's is akin to complete trimming. Note that $k$ is a tuning constant which must be chosen. We only consider `default' tuning parameters in this paper which have 95\% asymptotic efficiency under a normal distribution. More detailed results can be found in both \cite{huber2009} and \cite{maronna2006}. These values are $k=1.345$ for Huber's $\rho$ and $k=4.685$ for Tukey's.

For the $\chi$ function, \cite{huber1964} proposed the choice 
\begin{equation}
\label{HuberScale}
\chi(x)=
\begin{cases}
x^2-\delta, &\ |x|\leq k\\
k^2-\delta, &\ |x|> k .
\end{cases}
\end{equation}
For consistency under the normal distribution the tuning parameter is set to $\delta\approx 0.71016$, taking $k = 1.345$ as above for Huber's $\psi$ is
referred to as Huber's `Proposal 2'. This estimator is used throughout this paper.
